NeuronUnit: Collaboratively Testing the Validity of Neuroscientific Models

### 2013-11-30 NeurounitFrontiers/Paper.tex

Each of these examples has leveraged \emph{ad hoc} infrastructure to support model validation. 
While the specific criteria used to evaluate models can vary widely between modeling domains, the underlying methodology is common and could be implemented once. 
Recognizing this, and inspired by unit testing practices, we have developed a discipline-agnostic framework for developing \emph{model validation test suites} called \textit{SciUnit} \citep{omar_sciunit_2013} (available from \url{http://sciunit.scidash.org}) \todo{cite sciunit paper again here?}. 
In this paper, we will begin by detailing \textit{SciUnit}, focusing on examples from single-neuron physiology (Sec. \ref{sec:sciunit}).

\textit{SciUnit} contains validation logic common across scientific disciplines.
But a particular discipline, such as neurophysiology, might have more specialized logic associated with it that can be shared amongst its sub-disciplines (e.g. hippocampal neurophysiology). 
We anticipate a collaborative workflow where this common logic is developed in common repositories on a social coding service, here \textit{Github}. 
For neurophysiology, we have developed such a repository, called \textit{NeuronUnit}. 
This repository contains common testing logic as well as bridges to existing informatics infrastructure to make it easy to import existing data and models. 
In particular, we will show how models described using \text{NeuroML} and provided freely by the \textit{Open Source Brain Project} (OSB, \cite{gleeson_open_2012}, \url{http://www.opensourcebrain.org}) can be tested in a fully automated fashion using published data curated by the \textit{NeuroElectro Project} (NeuroElectro, \cite{tripathy_neuroelectro:_2012}, \url{http://neuroelectro.org}), leveraging facilities from the \textit{NeuroTools} library (\url{http://neuralensemble.org/NeuroTools}) to extract features from model outputs (Sec. \ref{sec:neuronunit}). 
Scientists in particular sub-disciplines, like hippocampal neurophysiologists, can use this common logic to select relevant models and tests, forming focused test suites in common repositories. 
The end result of this workflow is a table like the one shown in Figure \ref{fig:sciunit_overview}, where submitted models can be compared based on the scores they achieve on different selected tests. 
These tables serve as a summary of the status of a modeling community, just as the results from a suite of unit tests serve as a summary of the status of a software project, and help scientists answer the questions mentioned earlier in this section more easily and accurately.

@@

\subsection{A Collaborative Workflow}\label{sec:neuronunit_acitivities}
The \textit{NeuronUnit} package contains a collection of test families and associated capabilities  relevant to neurophysiology, such as the tests used in the QSNMC, and is collaboratively developed on Github (\url{http://github.com/scidash/neuronunit})\todo{Must change name of repository on github to neuronunit}. Scientists can use these tools to easily create test suites tailored to specific neural systems of interest. For example, scientists interested specifically in testing models of cells in the hippocampus would use \textit{NeuronUnit} to create tests in a Github repository called \textit{hippocampus\_tests},  parameterizing them by relevant data from \textit{NeuroElectro}. To make this easy, \textit{NeuronUnit} contains \emph{bridge functions} for retrieving data and generating tests from \textit{NeuroElectro} directly (Sec. \ref{sec:neuroelectro}).

@@

To submit a new model for testing, a scientist can similarly fork the \textit{hippocampus\_tests} repository, add a new model to the relevant suite using the simple \textit{NeuroML}-\textit{NeuronUnit} bridge described in Sec. \ref{sec:neuroml_models} and test the model against available tests locally. When satisfied with parameter choices, the scientist can submit the model to \textit{OSB} and use the common \textit{OSB-NeuronUnit} bridge, also described in Sec. \ref{sec:neuroml_models}, to refer to it, and finally submit a pull request to the common repository. Once a model is in the repository, it will be evaluated on an ongoing basis against the latest accepted collections of tests, which use accepted statistical best practices, without needing the participation of the original modeler, assuming that the capabilities that need to be exposed by models stay stable.

@@

The end result of these efforts will be tables, like the one shown in Figure \ref{fig:sciunit_overview}, that allow scientists to understand, precisely, the state-of-the-art in their field. We are developing a lightweight portal, \textit{SciDash}, that directs scientists to the canonical test repositories in their field in order to prevent fragmentation and encourage community participation \todo{cite icse paper?}. \textit{SciDash} will also facilitate viewing test results without needing to clone test repositories and run tests locally for exploratory purposes, based on the \textit{iPython notebook} technology mentioned in Sec. \ref{sec:visualization}. We leave details of \textit{SciDash} as future work, focusing this paper on the core testing methodology (Sec. 2) and on the details of the bridge technologies mentioned above.

>>>>

Each of these examples has leveraged \emph{ad hoc} infrastructure to support model validation. 
While the specific criteria used to evaluate models can vary widely between modeling domains, the underlying methodology is common and could be implemented once. 
Recognizing this, and inspired by unit testing practices, we have developed a discipline-agnostic framework for developing \emph{model validation test suites} called \textit{SciUnit} \citep{omar_sciunit_2013} (available from \url{http://sciunit.scidash.org}). 
In this paper, we will begin by detailing \textit{SciUnit}, focusing on examples from single-neuron physiology (Sec. \ref{sec:sciunit}).

\textit{SciUnit} contains validation logic common across scientific disciplines.
But a particular discipline, such as neurophysiology, might have more specialized logic associated with it that can be shared amongst its sub-disciplines (e.g. hippocampal neurophysiology). 
We anticipate a collaborative workflow where this common logic is developed in common repositories on a social coding service, here \textit{GitHub}. 
For neurophysiology, we have developed such a repository, called \textit{NeuronUnit}. 
This repository contains common testing logic as well as bridges to existing informatics infrastructure to make it easy to import existing data and models. 
In particular, we will show how models described using \text{NeuroML} and provided freely by the \textit{Open Source Brain Project} (OSB, \cite{gleeson_open_2012}, \url{http://www.opensourcebrain.org}) can be tested in a fully automated fashion using published data curated by the \textit{NeuroElectro Project} (NeuroElectro, \cite{tripathy_neuroelectro:_2012}, \url{http://neuroelectro.org}), leveraging facilities from the \textit{NeuroTools} library (\url{http://neuralensemble.org/NeuroTools}) to extract features from model outputs (Sec. \ref{sec:neuronunit}). 
Scientists in particular sub-disciplines, like hippocampal neurophysiologists, can use this common logic to select relevant models and tests, forming focused test suites in common repositories. 
The end result of this workflow is a table like the one shown in Figure \ref{fig:sciunit_overview}, where submitted models can be compared based on the scores they achieve on different selected tests. 
These tables serve as a summary of the status of a modeling community, just as the results from a suite of unit tests serve as a summary of the status of a software project, and help scientists answer the questions mentioned earlier in this section more easily and accurately.

@@

\subsection{A Collaborative Workflow}\label{sec:neuronunit_acitivities}
The \textit{NeuronUnit} package contains a collection of test families and associated capabilities relevant to neurophysiology, such as the tests used in the QSNMC, and is collaboratively developed on GitHub (\url{http://github.com/scidash/neuronunit}). 
Scientists can use these tools to easily create test suites tailored to specific neuronal systems of interest. 
For example, scientists interested specifically in testing models of cells in the hippocampus would use \textit{NeuronUnit} to create tests in a GitHub repository called \textit{hippocampus\_tests}, parameterizing them by relevant data from \textit{NeuroElectro}. 
To make this easy, \textit{NeuronUnit} contains \emph{bridge functions} for retrieving data and generating tests from \textit{NeuroElectro} directly (Sec. \ref{sec:neuroelectro}).

@@

To submit a new model for testing, a scientist can similarly fork the \textit{hippocampus\_tests} repository, add a new model to the relevant suite using the simple \textit{NeuroML}-\textit{NeuronUnit} bridge described in Sec. \ref{sec:neuroml_models} and test the model against available tests locally. 
When satisfied with parameter choices, the scientist can submit the model to \textit{OSB} and use the common \textit{OSB-NeuronUnit} bridge, also described in Sec. \ref{sec:neuroml_models}, to refer to it, and finally submit a pull request to the common repository. 
Once a model is in the repository, it will be evaluated on an ongoing basis against the latest accepted collections of tests, which use accepted statistical best practices, without needing the participation of the original modeler, assuming that the capabilities that need to be exposed by models stay stable.  
GitHub is an excellent platform for this kind of activity, and is becoming a force for scientific collaboration, reproducibility, and transparency \citep{ram_git_2013}.

@@

The end result of these efforts will be tables, like the one shown in Figure \ref{fig:sciunit_overview}, that allow scientists to understand, precisely, the state-of-the-art in their field. We are developing a lightweight portal, \textit{SciDash}, that directs scientists to the canonical test repositories in their field in order to prevent fragmentation and encourage community participation \citep{omar_sciunit_2013}. \textit{SciDash} will also facilitate viewing test results without needing to clone test repositories and run tests locally for exploratory purposes, based on the \textit{iPython notebook} technology mentioned in Sec. \ref{sec:visualization}. We leave details of \textit{SciDash} as future work, focusing this paper on the core testing methodology (Sec. 2) and on the details of the bridge technologies mentioned above.
